# Exercise 5

# Overview
```{r}
human <- read.csv('human.csv',row.names = 1)
str(human)
```
```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(GGally)
ggpairs(human)
cor(human) %>%corrplot::corrplot()
```

# PCA on the non standardized data
Now it is time to perform PCA. I am going to do it on the non standardized data.
```{r warning=FALSE}
pca_non_sta <- prcomp(human)
summary(pca_non_sta)
biplot(pca_non_sta, choices = 1:2)
```
The results we got don't make sense at all. The reason for it is that our data is not standardized.

# PCA on the standardized data
Let us do PCA on the standardized data this time.
```{r}
human <- scale(human) # standardizing
pca_sta <- prcomp(human)
summary(pca_sta)
biplot(pca_sta, choices = 1:2, cex = c(0.5, 1), col = c("pink", "black"))
```
The analysis on the standardized data produces more realstic and more useful results than the previous one did.

# Tea data

Now we are going to do MCA on the tea dataset. 
```{r warning=FALSE}
# download "tea"
library(FactoMineR)
library(tidyr)
data(tea)
tea <- dplyr::select(tea, one_of(c("tea.time", "tearoom", "Tea","age_Q", "sex")))
tea <- filter(tea, complete.cases(tea))
gather(tea) %>% ggplot(aes(value)) +
  facet_wrap("key", scales = "free") +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size= 9))
mca <- MCA(tea, graph = FALSE)
plot(mca, invisible = c("ind"))
```